---
title: "GSS_Run & Print OLS"
author: "Rachel Ganly"
date: "9/22/2021"
output:
  html_document: default
---

This document provides instructions and code for running regressions using GSS data for the CaND3 data replication exercise. 

1. Read in cleaned datasets, and load packages. 

```{r, include=FALSE}

#---- Clear data space
rm(list = ls())

# load packagaes

library(tidyverse)
library(ggplot2)
library(lmtest)
library(sandwich)
library(stargazer)
library(MASS)

# Read data

load(file="GSS_Health_NA.csv")
load(file="GSS_Health_Cats.csv")
#There are 20,602 rows and 1,316 columns

```

2. Fit OLS Regressions with standard and robust errors.

- Formula for regressions: Subjective Wellbeing ~ Self-rated health + self-rated mental health + family income + educational attainment + sex + age group + aboriginal status + ethnic minority status. 

```{r, include=FALSE}

#Fit using OLS
ols<-lm(SWB~SRH+MENTAL_SRH+INCOME+EDU+SEX+AGE+ABOR+ETHNIC_MIN, data=Health_NA)
#show classic s.e.
classic<-summary(ols)
classic


#calculate robust s.e. using coeftest
robust<-coeftest(ols, vcov.=vcovHC(ols,type="HC1"))
cov<-vcovHC(ols, type="HC")
robust.se<-sqrt(diag(cov))
robust

```

- Compare Standard Errors. 
```{r, include=FALSE}
#Compare the standard errors
cse<-classic$coefficients[,2]
rse<-robust[,2]
comparese<-cbind(cse,rse)
colnames(comparese)<-c("Classic standard error","Robust standard error")
comparese

```

3. Fit logistic model

- Create indicator variable for High Subjective Wellbeing whereby a score of 9 or above is "High" (1) and 8 or below is "Low" (0). 
- Fit a logit model. 
- Formula for logit: High Subjective Wellbeing ~ Self-rated health + self-rated mental health + family income + educational attainment + sex + age group + aboriginal status + ethnic minority status. 

```{r, include=FALSE}
#Create new indicator variable for High Subjective Wellbeing (9 or 10 out of 10)
Health_NA<-Health_NA%>%mutate(HighSWB=ifelse(is.na(SWB), NA, ifelse(SWB>=9,1,0)))
Health_Cats<-Health_Cats%>%mutate(HighSWB=ifelse(is.na(SWB), NA, ifelse(SWB==9,1, ifelse(SWB==10,1,0))))

#Check new variable
Health_NA%>%count(HighSWB,SWB)
Health_Cats%>%count(HighSWB,SWB)

# Run logit model
logit<-glm(HighSWB~SRH+MENTAL_SRH+INCOME+EDU+SEX+AGE+ABOR+ETHNIC_MIN, data=Health_NA, family = binomial(link = "logit"))
summary(logit)
```


4. Create Descriptive Data Table. 
- Load cleaned dataframe which contains factors for all categorical variables. 
- Recode labels for all relevant variables. 
- Drop variables for Subjective Wellbeing and Age Group which will not be shown in the descriptive data. 
- Create labels for descriptive data table. 

```{r, include=FALSE}

library(vtable)

#Select and mutate relevant variable for the table

DESCRIP<-Health_Cats%>%dplyr::select(-SWB,-AGE)%>%
  mutate(REGION=as.factor(REGION))%>%
  mutate(SEX=case_when(SEX==1 ~ "Male",SEX==2 ~ "Female"),
         INCOME=case_when(INCOME==1 | INCOME==2 ~ "Less than $50,000",
                          INCOME==3 | INCOME==4 ~ "$50,000 to $99,999",
                          INCOME==5 | INCOME==5 ~ "$100,000 or over"),
         ETHNIC_MIN=case_when(ETHNIC_MIN==1 ~ "Ethnic Minority",ETHNIC_MIN==2 ~ "Non-Ethnic Minority"),
         ABOR=case_when(ABOR==1 ~ "Aboriginal Heritage",ABOR==2 ~ "Non-Aboriginal Heritage"),
         EDU=case_when(EDU==1 | EDU==2 ~ "1. High School Diploma or Below",
                          EDU==3 | EDU==4 | EDU==5 ~ "2. Non-Degree Tertiary Education",
                          EDU==6 | EDU==7 ~ "3. Bachelor's Degree or Above"),
         REGION=case_when(REGION==1 ~ "Atlantic",REGION==2 ~ "Quebec",
                          REGION==3 ~ "Onatrio",REGION==4 ~ "Prarie",REGION==5 ~ "British Colombia"),
         SRH=case_when(SRH==1 ~ "1. Excellent",SRH==2 ~ "2. Very Good",
                          SRH==3 ~ "3. Good",SRH==4 ~ "4. Fair",SRH==5 ~ "5. Poor"),
         MENTAL_SRH=case_when(MENTAL_SRH==1 ~ "1. Excellent",MENTAL_SRH==2 ~ "2. Very Good",
                          MENTAL_SRH==3 ~ "3. Good",MENTAL_SRH==4 ~ "4. Fair",MENTAL_SRH==5 ~ "5. Poor"),
         Subjective_Wellbeing=case_when(HighSWB==1 ~ "High",HighSWB==0 ~ "Not High"))%>%
  dplyr::select(-HighSWB)

DESCRIP_LABS <- c("Sex", "Region", "Household Income", "Ethnic Minority Status", "Self-Rated Health", "Self-Rated Mental Health", "Educational Attainment", "Age (years)", "Aboriginal Heritage")


```

5. Print Descriptive Data Table. 
- Use vtable package to print summary statistics showing N and Mean/Proportion values. 

```{r, echo=FALSE}
#Direct Output

st(DESCRIP, group="Subjective_Wellbeing",title="Summary Statistics",  digits=2, summ=c('notNA(x)','mean(x)'), labels = DESCRIP_LABS, summ.names = c("Number of Individuals", "Mean"))

#Latex Code
st(DESCRIP, group="Subjective_Wellbeing",title="Summary Statistics",  digits=2, summ=c('notNA(x)','mean(x)'), labels = DESCRIP_LABS, summ.names = c("Number of Individuals", "Mean"), out="latex")

```


6. Set up Regression Results Table
- Create labels for results table. 

```{r, include=FALSE}
#Create regression table using Stargazer

stargazer(ols,ols,logit, se=list(NULL, robust.se, NULL), type="text", style="demography")

mylabs<- c("Poor Self-Rated Health", "Poor Self-Rated Mental Health","Income", "Education Level", "Sex","Age", "Aboriginal Origin", "Non-Aboriginal Ethnic Minority", "Constant")

```

7. Print Final Regression Results Table. 

- Use stargazer package to print regression results table showing 3 models; two OLS models with standard and robust errors respectively, and logit model. 
- Print directly and also print latex code. 

```{r, echo=FALSE}

#Direct Ouput

stargazer(ols,ols,logit, se=list(NULL, robust.se, NULL), column.sep.width="3pt", column.labels = c("SE", "Robust SE", ""),model.numbers = FALSE, colnames=FALSE, dep.var.labels= c("Wellbeing","High Wellbeing"), title="Association between Subjective Wellbeing and Demographic and Health Factors", align=TRUE, no.space=TRUE,
              style = "demography", covariate.labels=mylabs, type="text")

#Latex Code

stargazer(ols,ols,logit, se=list(NULL, robust.se, NULL), column.sep.width="3pt", column.labels = c("SE", "Robust SE", ""),model.numbers = FALSE, colnames=FALSE, dep.var.labels= c("Wellbeing","High Wellbeing"), title="Association between Subjective Wellbeing and Demographic and Health Factors", align=TRUE, no.space=TRUE,
              style = "demography", covariate.labels=mylabs, type="latex")

```

