---
title: "GSS_Data_Clean"
author: "Rachel Ganly"
date: "9/21/2021"
output: html_document
---

This document provides instructions and code for cleaning GSS data to be used in regressions for the CaND3 data replication exercise. 

1. Load required packages.

2. Read in GSS file named "gss-12M0025-E-2017-c-31_F1.csv". 

```{r, include=FALSE}

#---- Clear data space
rm(list = ls())

# load packagaes

library(tidyverse)
library(ggplot2)
library(reshape2)
library(lmtest)


# Read data

GSS<-read.csv("gss-12M0025-E-2017-c-31_F1.csv")
#There are 20,602 rows and 1,316 columns


```

3. Explore data set and check for missing values. 

```{r, include=FALSE}
ncol(GSS)
nrow(GSS)
sum(is.na(GSS)) 

GSS%>%count(SRH_110) # self rated health 1 excellent to 5 poor, 7 to 9 = missing
GSS%>%count(SRH_115) #self rated mental health 1 excellent to 5 poor, 7 to 9 = missing
GSS%>%count(SLM_01) #subjective wellbeing 0 very dissatisfied to 10 very satisfied, missing values 97 to 99 = dont know, refused, NA
GSS%>%count(EHG3_01B) #completed education #missing values 97 to 99 = dont know, refused, NA
GSS%>%count(AMB_01) # aboriginal respondent or not #missing value 6 = valid skip as not born in relevant country, 7-9 don't know/refused
GSS%>%count(AGEGR10) #age group 1=15-24 yrs, 7 = over 75 yrs #no missing
GSS%>%count(SEX) #sex  1= male, 2 = female #no missing
GSS%>%count(REGION) #region living # 5 regions, no missing
GSS%>%count(FAMINCG2) #family income 1 = <$25k, 6 = >$125k, no missing
GSS%>%count(VISMIN) #minority 1 = minority, 2 = not minority, 6 to 9 = missing


```

4. Clean data. 

- Select key variables for age group (by 10 years), sex, region, family income, aboriginal heritage, visible minority status, self-rated health, self-rated mental health, subjective well being, educational attainment and age in years

- Recode aboriginal variable so that those who do not originate from relevant countries (valid skip) are classed as non-aboriginal 

```{r, include=FALSE}
#Select relevant variables and rename
Health<-GSS%>%select(AGEGR10,SEX,REGION, FAMINCG2,AMB_01,VISMIN, SRH_110,SRH_115, SLM_01, EHG3_01B, AGEC)%>%rename(
  EDU=EHG3_01B,
  SRH=SRH_110,
  MENTAL_SRH=SRH_115,
  SWB=SLM_01,
  AGE=AGEGR10,
  INCOME=FAMINCG2,
  ABOR=AMB_01,
  ETHNIC_MIN=VISMIN)

ncol(Health)
colnames(Health)
nrow(Health)

#Recode 6 as 2 in ABOR
Health<-Health%>%mutate(ABOR_final=ifelse(ABOR==6,2,ABOR))%>%select(-ABOR)
Health<-Health%>%rename(ABOR=ABOR_final)
colnames(Health)
Health%>%count(ABOR)


save(Health, file = "GSS_Health.csv")

```
  
5. Deal with Missing Data
- Recode all variables' missing values to NA
- Create two dataframes: one which retains integer values for covariates, and another which converts relevant categorical variables in GSS to factors

```{r, include=FALSE}

load("GSS_Health.csv")
# Deal with missing data
Health_NA<-Health%>%mutate(
  EDU=ifelse(EDU>7,NA,EDU),
  SRH=ifelse(SRH>5,NA,SRH),
  MENTAL_SRH=ifelse(MENTAL_SRH>5,NA,MENTAL_SRH),
  SWB=ifelse(SWB>10,NA,SWB),
  AGE=ifelse(AGE>7,NA,AGE),
  INCOME=ifelse(INCOME>6,NA,INCOME),
  ABOR=ifelse(ABOR>=7,NA,ABOR),
  ETHNIC_MIN=ifelse(ETHNIC_MIN>=7,NA,ETHNIC_MIN),
SEX=ifelse(SEX>2, NA, SEX))

#Check NA values disappeared
glimpse(Health_NA)
sum(is.na(Health_NA))
Health_NA%>%count(EDU)
Health_NA%>%count(SRH)
Health_NA%>%count(MENTAL_SRH)
Health_NA%>%count(SWB) 
Health_NA%>%count(AGE)
Health_NA%>%count(INCOME)
Health_NA%>%count(ABOR)
Health_NA%>%count(ETHNIC_MIN)
Health_NA%>%count(SEX)

#Save cleaned file

save(Health_NA, file = "GSS_Health_NA.csv")

# Create df with categories (factors) rather than integers
Health_Cats<-Health_NA%>%mutate(
  EDU=as.factor(EDU),
  SRH=as.factor(SRH),
  MENTAL_SRH=as.factor(MENTAL_SRH),
  SWB=as.factor(SWB),
  AGE=as.factor(AGE),
  INCOME=as.factor(INCOME),
  ABOR=as.factor(ABOR),
  ETHNIC_MIN=as.factor(ETHNIC_MIN),
SEX=as.factor(SEX),
REGION=as.factor(REGION))


save(Health_Cats, file = "GSS_Health_Cats.csv")

```



